---
title: "heartbeat data alanysis - 1"
author: "km"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library activation
```{r, warning=F, error=F, message=F}
library("seewave")    # spectro, timer, specprop
library("tuneR")      # raedMP3, writeWave
library("dplyr")      # dataframe handling
library("ggplot2")    # make plot
```

# dataset
HeartSound & Murmur Library @ Univ. Michigan (.mp3)  
http://www.med.umich.edu/lrc/psb_open/html/repo/primer_heartsound/primer_heartsound.html

## inport data


```{r}
# file list in data folder
A <- list.files("Michigan_Heart_Sounds")

k <- 1
A[k]

# read file
dat <- readMP3(paste("Michigan_Heart_Sounds/", A[k], sep=""))

# data structure
str(dat)
```

# in partial data (first 1.6sec)
```{r, fig.width=8, fig.height=3}
start <- 0 * dat@samp.rate
end <- 1.6 * dat@samp.rate

dat1 <- dat[start:end]

# amplitude
par(mar = c(3, 5, 0.5, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
plot(dat1)

# spectrogram
spectro(dat1, 
        collevels = seq(-50, 0, length=20),  # dB range
        palette = reverse.gray.colors.1,     # color palette
        flim = c(0, 1))                      # kHz range
```

## start & end timing of the signals
- the function `timer` in {seewave} package computes and shows the duration of signal periods, pause periods and their ratio.  
- see detail: `help(timer)`
```{r, fig.width=8, fig.height=3}
# cutoff: 4.5% of signal
threshold <- 4.5

par(mar = c(3, 3, 0.5, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
dat_hb <- timer(dat1, threshold = threshold, msmooth = c(50, 0))

# data structure
str(dat_hb)
```
```{r, fig.width=8, fig.height=3, echo=F}
par(mar = c(1.5, 1.5, 1, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
hist(dat_hb$p,breaks = 50, main = "histgram of silent dulation: dat_hb$p")
```
  
## start and end timing of signals
```{r, fig.width=8, fig.height=3}
# ignor first and last silence
dat_hb$p[1] <- dat_hb$p[1]+1
dat_hb$p[length(dat_hb$p)] <- dat_hb$p[length(dat_hb$p)]+1

# ignor silent dulation < 0.275 sec
threshold_silent <- 0.275
a <- c(1:length(dat_hb$p))[dat_hb$p > threshold_silent]

# start and end timing of signals
list_start <- dat_hb$s.start[a]
list_end <- dat_hb$s.end[a-1]

```

## plot: signal timing
```{r, fig.width=8, fig.height=3, echo=T}
x <- 1:length(dat1@left)/dat@samp.rate
y <- x

# cutin
for(ii in 1:length(list_end)){
  y[y >= list_start[ii] & y <= list_end[ii]] <- 100
}
# cutoff
y[y < 100] <- 0

# plot
par(mar = c(3, 3, 0.5, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
plot(x, y, type = "l", col = "Red", yaxt = "n")
par(new = T)
plot(dat1)
```

- first signal  
```{r, fig.width=6, fig.height=8, out.width=200, echo=F}
List <- list.files(paste("wav/", A[1], sep=""))
dati <- readWave(paste("wav/", A[1], "/", List[1], sep=""))

par(mar = c(1.5, 1, 0.5, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
spectro(dati, 
        collevels = seq(-50, 0, length=20),  # dB range
        palette = reverse.gray.colors.1,     # color palette
        flim = c(0, 1), scale =F, osc =T)                      # kHz range
```
- 2nd signal  
```{r, fig.width=6, fig.height=8, out.width=200, echo=F}
List <- list.files(paste("wav/", A[1], sep=""))
dati <- readWave(paste("wav/", A[1], "/", List[2], sep=""))

par(mar = c(1.5, 1, 0.5, 0.5), tcl = -0.2, mgp = c(1.5, 0.3, 0))
spectro(dati, 
        collevels = seq(-50, 0, length=20),  # dB range
        palette = reverse.gray.colors.1,     # color palette
        flim = c(0, 1), scale =F, osc =T)                      # kHz range
```


# in whole data
## signal timing
```{r, fig.width=6, fig.height=3, warning=F}
# Whole data
dat_hb <- timer(dat, threshold = threshold, msmooth = c(50, 0), plot=F)

dat_hb$p[1] <- dat_hb$p[1]+1
dat_hb$p[length(dat_hb$p)] <- dat_hb$p[length(dat_hb$p)]+1

a <- c(1:length(dat_hb$p))[dat_hb$p > threshold_silent]

# start and end timing of signals
list_start <- dat_hb$s.start[a] * dat@samp.rate
list_end <- dat_hb$s.end[a-1] * dat@samp.rate
```

## split data
```{r, fig.width=6, fig.height=3, warning=F}
# make folder
 dir.create(paste("wav/", A[k], sep=""))

# calcurate parameters and save .wav of each signal
spec_call <- NULL
for(i in 1:length(list_end)){
  dat_i <- dat[list_start[i]:list_end[i]]
  
  spec_dat <- specprop(spec(dat_i, f = dat_i@samp.rate, plot =F))
  spec_dat$dulation <- length(dat_i@left)/dat_i@samp.rate
  spec_call[[i]] <- spec_dat
 writeWave(dat_i, paste("wav/", A[k], "/", i, "_",
                        substr(A[k],1, nchar(A[k])-3), "wav",sep=""))
}
```
## collect parameters 
- see detail: `help(specprop)`

parameters||
----|----|
mean|mean frequency (see mean)|
sd|  standard deviation of the mean (see sd)|
sem|  standard error of the mean|
median|  median frequency (see median)|
mode|  mode frequency, i.e. the dominant frequency|
Q25|  first quartile (see quantile)|
Q75|  third quartile (see quantile)|
IQR|  interquartile range (see IQR)|
cent|  centroid, see note|
skewness|  skewness, a measure of asymmetry, see note|
kurtosis|  kurtosis, a measure of peakedness, see note|
sfm|  spectral flatness measure (see sfm)|
sh|  spectral entropy (see sh)|
prec|  frequency precision of the spectrum|


```{r, fig.width=6, fig.height=3, warning=F, echo=F}
spec_summary <- data.frame(
  mean = sapply(spec_call, "[[", "mean"),
  sd = sapply(spec_call, "[[", "sd"),
  median = sapply(spec_call, "[[", "median"),
  sem = sapply(spec_call, "[[", "sem"),
  mode = sapply(spec_call, "[[", "mode"),
  Q25 = sapply(spec_call, "[[", "Q25"),
  Q75 = sapply(spec_call, "[[", "Q75"),
  IQR = sapply(spec_call, "[[", "IQR"),
  cent = sapply(spec_call, "[[", "cent"),
  skewness = sapply(spec_call, "[[", "skewness"),
  kurtosis = sapply(spec_call, "[[", "kurtosis"),
  sfm = sapply(spec_call, "[[", "sfm"),
  sh = sapply(spec_call, "[[", "sh"),
  prec = sapply(spec_call, "[[", "prec"),
  dulation = sapply(spec_call, "[[", "dulation")
)
```

## plot: distributions of heart beat signal parameters
- each point reprezent each signal in one .wav file
```{r, fig.width=8, fig.height=8, echo = F}
plot(spec_summary[,c("mean", "sd", "mode", "kurtosis",
                     "sfm", "sh", "dulation")], pch =20, cex =0.2)
```

# session information
```{r}
sessionInfo()
```